A search for high-mass black holes in data from the second LIGO observing run using a Bayesian ranking-statistic

### 2021-04-15 main.tex
edits for v2

\documentclass[%
%reprint,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
% preprint,
%preprintnumbers,
 nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps,
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
 twocolumn
]{revtex4-2}

@@

% Keywords
\newcommand{\bilby}{{\sc \href{https://lscsoft.docs.ligo.org/bilby/}{\texttt{Bilby}}}\xspace}
\newcommand{\bilbypipe}{{\sc bilby\_pipe}\xspace}
\newcommand{\pbilby}{{\sc pBilby}\xspace}
\newcommand{\dynesty}{{\sc dynesty}\xspace}
\newcommand{\cpnest}{{\sc cpnest}\xspace}
\newcommand{\ptemcee}{{\sc ptemcee}\xspace}
\newcommand{\gwpy}{{\sc \href{https://gwpy.github.io/}{\texttt{GWpy}}}\xspace}
\newcommand{\imrphenomp}{{\sc IMRPhenomPv2}\xspace}
\newcommand{\seob}{{\sc SEOBNRv4PHM}\xspace}
\newcommand{\gstlal}{{\sc GstLAL}\xspace}
\newcommand{\pycbc}{{\sc \href{https://pycbc.org/}{\texttt{PyCBC}}}\xspace}

@@

% author comments
\newcommand{\avi}[1]{\textcolor{orange}{[AV: #1]}}
\newcommand{\rs}[1]{\textcolor{red}{[RS: #1]}}
\newcommand{\et}[1]{\textcolor{blue}{[ET: #1]}}

% %%% AFFILIATIONS
% %------ affiliation shortcuts
% \newcommand{\SPAno}{1}
% \newcommand{\OzGravMonashno}{2}
% \newcommand{\CITno}{3}
% \newcommand{\MITno}{4}
% \newcommand{\Kavlino}{5}

% %--------------------------
% \newcommand{\SPA}{School of Physics and Astronomy, Monash University, Clayton VIC 3800, Australia}
% \newcommand{\OzGravMonash}{OzGrav: The ARC Centre of Excellence for Gravitational Wave Discovery, Clayton VIC 3800, Australia}
% \newcommand{\MIT}{LIGO Laboratory, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
% \newcommand{\Kavli}{Department of Physics and Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, \\ 77 Massachusetts Ave, Cambridge, MA 02139, USA}
% \newcommand{\CIT}{LIGO Laboratory, California Institute of Technology, Pasadena, CA 91125, USA}
% %%%

@@

\title{A Bayesian ranking statistic to find high-mass black holes in LIGO--Virgo data}

% \input{authors}
%\author[1]{Test}
%\affil[1]{Test Affil}

@@

\begin{abstract}
The detection of intermediate-mass black holes ( $10^2-10^6$~\msun) will shed light on the formation of supermassive black holes and thus galaxy formation. Although LIGO is sensitive to the merger of binary black holes with total masses up to $400$~\msun, only 4 of their 50 detections have a total mass $>100$~\msun with >95\% credibility. A possible explanation for the absence of intermediate-mass events may be their misclassification as short-duration instrumental noise transients. Short-duration instrumental transients mimic the short-duration gravitational-wave signals from intermediate-mass binary black hole mergers. 
\et{I think the preceding text is misleading. We make it sound like we expect a uniform distribution of total mass, and so it's a surprising that we don't see IMBH. However, in reality, it is theoretically challenging to make IMBH. You need an exceptionally massive progenitor star to avoid pair instability supernova. Or, maybe you can make IMBH from hierarchical mergers or accretion. No one expects them to be common, but they are  interesting because of their connection to supermassive black holes. I suggest rewriting the preceding text to explain how IMBH might form and why they are interesting to look for.}
Here we demonstrate a ranking statistic utilising Bayesian inference for the detection of high-mass binary black hole mergers (with a total mass $>55$~\msun). \et{Claims of beating matched filtering not yet borne out.} We apply this technique on the high-mass triggers during LIGO's second observing run to search for previously unresolved gravitational-wave signals from high-mass binary black holes. Our analysis does not discover new gravitational-wave events. However, we find support for some previously identified borderline detections.
\et{Don't oversell Bayesian measure since our method is only semi-Bayesian.}
\end{abstract}

@@

%%%%%%---SECTIONS---%%%%%%%%%%%%
\section{Introduction}
Since the 1970s, there has been a steady accumulation of evidence for stellar mass ($\text{M}_\text{BH} < 10^{2} \ \msun$) and supermassive black holes ($\text{M}_\text{BH} > 10^{5} \ \msun$)~\cite{Webster:1972:Natur, Balick:1974:ApJ, Ghez:1998:ApJ, Genzel:2010:RvMP, Abbott:2019:PhRvX, EventHorizonTelescopeCollaboration:2019:ApJL, Abbott:2020:arXiv}. However, there is a deficiency of observational evidence for black holes in the intermediate-mass range $10^{2} - 10^{5}\ \text{M}_{\odot}$. Astronomers have detected a handful of intermediate-masses candidates using reverberation mapping, taking direct kinematic measurements, applying  M-$\sigma$ and M-L relations, and also studying  X-ray luminosity and spectra~\cite{Greene:2004:ApJ, Graham:2013:ApJ, Mezcua:2017:IJMPD, Koliopanos:2017:mbhe, Lin:2020:ApJL, Greene:2020:ARA&A}. However, these observations are highly uncertain and can be attributed to sources without intermediate-mass black holes~\cite{kiziltan2017, freire2017}. In the last few years, astronomers have found definitive evidence for two intermediate-mass black holes, (i) the $142^{+28}_{-16}\ \msun$ remnant observed from the gravitational wave event GW190521~\cite{Abbott:2020:PhRvL}, and the $5.5^{+1.7}_{-0.9}\times10^{4}\ \msun$ black hole identified from gravitational lensing in the light curve of GRB950830~\cite{paynter_evidence_2021}. The concrete discovery of more intermediate-mass black holes will bridge the observational gap and illuminate our understanding of galaxy and supermassive black hole formation. Additionally, a catalog of observed intermediate-mass black holes can act as probes to their formation environments, such as in the accretion disks of active galactic nuclei~\cite{}, the centers of dense stellar clusters~\cite{}, and even Population-III stars~\cite{}.

% Given the existence of supermassive black holes detected in the early universe, it is likely for intermediate-mass black holes to have also existed~\cite{Banados:2018:Natur}. However, we only have a handful of promising candidates~\cite{Greene:2004:ApJ, Graham:2013:ApJ, Mezcua:2017:IJMPD, Koliopanos:2017:mbhe, Lin:2020:ApJL, Greene:2020:ARA&A}, and two confirmed intermediate-mass black holes, (i) the $142^{+28}_{-16}\ \msun$ remnant observed from the gravitational wave event GW190521~\cite{Abbott:2020:PhRvL}, and the $5.5^{+1.7}_{-0.9}\times10^{4}\ \msun$ black hole identified from gravitational lensing in the light curve of GRB950830~\cite{paynter_evidence_2021}. The concrete discovery of more intermediate-mass black holes will bridge the observational gap and illuminate our understanding of galaxy and supermassive black hole formation. Additionally, a catalog of observed intermediate-mass black holes can act as probes to their formation environments, such as in the accretion disks of active galactic nuclei~\cite{}, the centers of dense stellar clusters~\cite{}, and even Population-III stars~\cite{}.

>>>>

\documentclass[%
%reprint,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
% preprint,
%preprintnumbers,
 nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps,
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
 twocolumn,
 superscriptaddress
]{revtex4-2}

@@

% Keywords
\newcommand{\bilby}{{\sc \href{https://lscsoft.docs.ligo.org/bilby/}{\texttt{Bilby}}}\xspace}
\newcommand{\bilbypipe}{{\sc bilby\_pipe}\xspace}
\newcommand{\pbilby}{{\sc pBilby}\xspace}
\newcommand{\dynesty}{{\sc dynesty}\xspace}
\newcommand{\cpnest}{{\sc cpnest}\xspace}
\newcommand{\ptemcee}{{\sc ptemcee}\xspace}
\newcommand{\gwpy}{{\sc \href{https://gwpy.github.io/}{\texttt{GWpy}}}\xspace}
\newcommand{\imrphenomp}{{\sc IMRPhenomPv2}\xspace}
\newcommand{\seob}{{\sc SEOBNRv4PHM}\xspace}
\newcommand{\gstlal}{{\sc GstLAL}\xspace}
\newcommand{\cwb}{{\sc cWB}\xspace}
\newcommand{\pycbc}{{\sc \href{https://pycbc.org/}{\texttt{PyCBC}}}\xspace}

\newcommand{\GWTC}{{\sc \href{https://ui.adsabs.harvard.edu/abs/2019PhRvX...9c1040A/abstract}{\texttt{GWTC-1}}}\xspace}
\newcommand{\OGC}{{\sc \href{https://ui.adsabs.harvard.edu/abs/2020ApJ...891..123N/abstract}{\texttt{2-OGC}}}\xspace}
\newcommand{\IAS}{{\sc \href{https://ui.adsabs.harvard.edu/abs/2020PhRvD.101h3030V/abstract}{\texttt{IAS}}}\xspace}

@@

% author comments
\newcommand{\avi}[1]{\textcolor{orange}{[AV: #1]}}
\newcommand{\greg}[1]{\textcolor{purple}{[Greg: #1]}}
\newcommand{\rs}[1]{\textcolor{red}{[RS: #1]}}
\newcommand{\et}[1]{\textcolor{blue}{[ET: #1]}}

@@

\title{A search for high-mass black holes in data from the second LIGO observing run using a Bayesian ranking-statistic}

\input{authors}

@@

\begin{abstract}
The detection of an intermediate-mass black hole population ($10^2-10^6$~\msun) will provide probes to their formation environments (e.g., disks of active galactic nuclei, globular clusters) and illuminate a potential pathway to produce supermassive black holes. Numerous methods to search for $10^4-10^6\ \msun$ intermediate-mass black holes exist, while the primary method to find $10^2-10^3\ \msun$ is by identifying gravitational-waves emitted from the merger of stellar binaries. Theoretically, ground-based gravitational-wave detectors are sensitive to such mergers and have found one $142^{+28}_{-16}\ \msun$ intermediate-mass black hole. In reality, ground-based detectors are plagued with short-duration instrumental transients that mimic the gravitational-wave signals from mergers capable of producing intermediate-mass black holes. Here we demonstrate a ranking statistic utilizing Bayesian inference to detect high-mass binary black hole mergers (with a total mass $>55$~\msun). We apply this technique to the high-mass triggers during LIGO's second observing run to search for previously unresolved gravitational-wave signals from high-mass binary black holes. Our analysis does not discover new intermediate-mass black holes. However, we find support for some previously identified borderline stellar-mass binary black hole detections and gravitational-wave detections made externally from the LIGO-Virgo-KAGRA collaboration.
\end{abstract}

@@

%%%%%%---SECTIONS---%%%%%%%%%%%%
\section{Introduction}
Since the 1970s, there has been a steady accumulation of evidence for stellar mass ($\text{M}_\text{BH} < 10^{2} \ \msun$) and supermassive black holes ($\text{M}_\text{BH} > 10^{6} \ \msun$)~\cite{Webster:1972:Natur, Balick:1974:ApJ, Ghez:1998:ApJ, Genzel:2010:RvMP, Abbott:2019:PhRvX, EventHorizonTelescopeCollaboration:2019:ApJL, Abbott:2020:arXiv}.  However, there is a deficiency of observational evidence for black holes in the intermediate-mass range $10^{2} - 10^{6}\ \msun$. The discovery of an intermediate-mass black hole population will bridge this observational gap, probe intermediate-mass formation environments (e.g. accretion disks of active galactic nuclei~\cite{Tagawa:2021:ApJ, Li:2021:arXiv, Samsing:2020:arXiv, Tagawa:2020:ApJ, Ishibashi:2020:A&A, Grobner:2020:A&A, Yang:2019:PhRvL, McKernan:2019:ApJL, Yang:2019:ApJ, McKernan:2018:ApJ, Bellovary:2016:ApJL, McKernan:2014:MNRAS, McKernan:2012:MNRAS}, the centers of dense stellar clusters~\cite{Banerjee:2021:MNRASa, Zevin:2021:ApJ,Mapelli:2021:arXiv,Weatherford:2021:ApJL, Bouffanais:2021:arXiv, Ballone:2021:MNRAS, Kumamoto:2021:arXiv, Banerjee:2021:MNRASb, Martinez:2020:ApJ, Romero-Shaw:2020:ApJL, Anagnostou:2020:PASA}, Population-III stars~\cite{Toubiana:2021:PhRvL, Farrell:2021:MNRAS, Safarzadeh:2020:ApJL, Liu:2020:MNRAS, Inayoshi:2017:MNRAS}), and illuminate our understanding of supermassive black hole formation~\cite{Askar:2021:MNRAS, ArcaSedda:2019:arXiv, Amaro-Seoane:2007:CQGra, Gurkan:2006:ApJL}. A variety of techniques have been employed to search for $10^{4} - 10^{6}\ \msun$ intermediate mass candidates (reverberation mapping~\cite{Peterson:2014:SSRv}, direct kinematic measurements~\cite{Schodel:2002:Natur, Kiziltan:2017:Natur}, applying macroscopic galaxy to black hole mass scaling relations (M$_{BH}$-$\sigma$ and M$_{BH}$-L relations)~\cite{Graham:2013:ApJ, Wevers:2017:MNRAS}, studying  X-ray luminosity and spectra~\cite{Greene:2004:ApJ, Lin:2020:ApJL}, gravitational lensing of light curves~\cite{paynter_evidence_2021}, and others~\cite{Greene:2020:ARA&A, Koliopanos:2017:mbhe, Mezcua:2017:IJMPD})). However, these observational techniques are challenging to use for intermediate mass black holes, due to their small sphere of influence. Additionally, some candidates discovered with these techniques can be attributed to sources without intermediate-mass black holes (e.g. clusters of stellar-mass black holes~\cite{Ridolfi:2016:MNRAS, Freire:2017:MNRAS}, anisotropic emission from neutron stars~\cite{Israel:2017:MNRAS, RodriguezCastillo:2020:ApJ}), while others have high uncertainties due to the observational techniques utilized~\cite{Greene:2020:ARA&A}. Compact binaries coalesces (CBCs) can provide unambiguous gravitational wave signals for intermediate-mass candidates (e.g. the $142^{+28}_{-16}\ \msun$ remnant observed from the gravitational wave event GW190521~\cite{Abbott:2020:PhRvL}), and ground-based detectors can probe the lower-end $10^{2} - 10^{3}\ \msun$ of the intermediate mass range. Mergers involving either a binary system with at least one intermediate-mass component, or a system whose merger results in an intermediate-mass remnant will have large total masses $\text{M}_{T}$. As a binary's $\text{M}_{T}$ is associated with its gravitational-wave merger frequency, $f\sim \text{M}_{T}^{-1}$, systems with large $\text{M}_{T}$ have very low merger frequencies $f < 100\ \text{Hz}$. Hence, ground based gravitational wave detectors ($\sim 10^1 - 10^3\ \text{Hz}$) are sensitive to the last milliseconds of merging systems with $100 \msun < \text{M}_{T} < 400 \msun$~\cite{LIGOScientificCollaboration:2015:CQGra, Martynov:2016:PhRvD, Moore_2014}, while space-based detectors ($\sim 10^{-2} - 10^1\ \text{Hz}$) can study the full signals of merging systems with $10^4 \msun < \text{M}_{T} < 10^7 \msun$~\cite{ Moore_2014, Lu:2019:PhRvD}.

Because of high-total mass systems' short-durations in ground-based gravitational wave data steams, handling data quality is critical to their detection. The low-frequency ranges of ground-based observatories are plagued with incoherent non-stationary terrestrial artifacts, called \textit{glitches}~\cite{ pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. Some glitches, similar to signals from high total mass mergers, last for a fraction of a second, making them difficult to distinguish from the signals. These glitches that mimic astrophysical signals can severely decrease the confidence in detection of true gravitational-waves from high total mass mergers.

Although a significant fraction of the glitches can be removed by testing them for coherence amongst various ground based detectors and performing matched-filtering, these methods are insufficient to remove all the glitches~\cite{ pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. One method to account for more glitches while searching for high total mass CBC gravitational-waves is by utilizing a Bayesian odds~\cite{bci, kanner2016leveraging, BCR1, BCR2, bcr_gw151216, bayesian_odds}. In this paper, we utilize a Bayesian method, called the Bayesian Coherence Odds-Ratio \bcr~\cite{BCR1}, to rank the candidate gravitational-wave signals from high-mass compact binary coalescences (systems with total masses in the range of $55-500\ \msun$) in the detector data recorded during O2. The \bcr used in this study is a bootstrap-Bayesian odds computed using Bayesian evidences that describe the explicit probability of data under the hypothesis that the data contains coherent signals vs incoherent glitches, and empirically calculated prior-odds that describe how likely each hypotheses is. The high-mass candidate \bcr values are then used to calculate the probability that the candidate is of astrophysical origins, \pastrobcr. Finally, if the candidate has a \pastro reported by the the LIGO-Virgo-KAGRA (LVK) collaboration in \GWTC~\cite{GWTC1}, the \pycbc-team~\cite{pycbc_code, pycbc_og0, pycbc_og1, pycbc_og2, pycbc_og3, pycbc_og4, pycbc_og5, pycbc_og6, pycbc_single_det, pycbc_ogc_2}, by the Institute of Advanced study's team (IAS)~\cite{IAS0, IAS1, IAS2}, or by \citet{bayesian_odds}, we compare their \pastro with the \pastrobcr.

We find that (a) high-mass events reported in the \GWTC, including GW170729 (the heaviest event in \GWTC) are very statistically significant $\pastrobcr>0.9$; (b) three out of the eight \IAS events and candidates have $\pastrobcr>0.5$, corroborating \IAS's detection claims for those events (GW170304, GW170727, GW170817A), and that (c) our ranking statistic does not identify any new intermediate-mass black holes, but does identify an unreported marginal stellar-mass binary black hole candidate, 170222 with $\pastrobcr\sim0.5$.

The remainder of this paper is structured as follows. We outline our methods, including details of our ranking statistic and the retrieval of our candidate in Section~\ref{sec:method}. We present details on the implementation of our analysis in Section~\ref{sec:Analysis}. Finally, we present our results in Section~\ref{sec:Results}, and discuss these results in the context of the significance of gravitational-wave candidates in Section~\ref{sec:Conclusion}.

The standard framework to identify CBC gravitational-wave signals hidden in data is by quantifying the significance of candidates with null-hypothesis significance testing~\cite{GWTC1, GWTC2}. In this framework, the candidates' ranking statistic is compared against a background distribution in a frequentist approach. The independent matched-filter searches, e.g. \pycbc~\cite{pycbc_og4} and \gstlal~\cite{sachdev2019gstlal}, and the coherent burst search \cwb~\cite{cwb} used by LVK to search for signals in gravitational-wave data all use ranking statistics in such a manner~\cite{GWTC1}. Both \pycbc and \gstlal's ranking statistic incorporate information of the relative likelihood that the data contains a coherent-signal versus just noise, while \cwb's ranking statistic uses the information of coherent energy present in the network of detectors~\cite{GWTC1}.

In a similar vein, this work utilizes Bayesian inference to calculate the Bayesian Coherence Odds-Ratio~\cite{BCR1}, \bcr, of high-mass candidates in LIGO's second observing run. We use the \bcr not as an odds-ratio but instead as a ranking statistic.

\subsection{The Bayesian Coherence Odds-Ratio}

Bayes theorem states that the posterior probability distribution $p(\parameters|d,\mathcal{H})$ for data $d$ and a vector of parameters \parameters that describe a model which quantifies a hypothesis $\mathcal{H}$, is given by
\begin{equation}
p(\parameters|d,\mathcal{H}) = \frac{\mathcal{L}(d|\parameters, \mathcal{H}) \ \pi(\parameters | \mathcal{H})}{\mathcal{Z}(d|\mathcal{H})}\ , 
\end{equation}
where $\mathcal{L}(d|\parameters, \mathcal{H})$ is the likelihood of the data given the parameters \parameters and the hypothesis, $\pi(\parameters | \mathcal{H})$ is the prior probability of the parameters, and finally,
\begin{equation}
    \mathcal{Z}(d|\mathcal{H}) = \int\limits_{\parameters} \ \mathcal{L}(d|\parameters,\mathcal{H}) \ \pi(\parameters | \mathcal{H}) \ d\parameters
\end{equation} is the likelihood after marginalizing over the parameters \parameters.  To compare two hypotheses $\mathcal{H}_A$ and $\mathcal{H}_B$ with the Bayes theorem one can calculate an odds-ratio
\begin{equation}
    \mathcal{O}^A_B = \frac{\mathcal{Z}^A\ \pi(\parameters | \mathcal{H_A})}{\mathcal{Z}^B\ \pi(\parameters | \mathcal{H_B})}\ ,
\end{equation}
where $\{\mathcal{Z}^A, \mathcal{Z}^B\}$ are shorthand for the evidences $\{\mathcal{Z}(d|\mathcal{H}_A), \mathcal{Z}(d|\mathcal{H}_B)\}$ and $\{\pi(\parameters | \mathcal{H_A}), \pi(\parameters | \mathcal{H_B})\}$ are the prior-odds for each hypotheses. The odds-ratio can tell us which of the two hypotheses is more likely. For example, if $\mathcal{O}^A_B >> 1$, then this odds ratio indicates that the $\mathcal{H}_A$ describes the data much better than $\mathcal{H}_B$.

The \bcr is a Bayesian odds-ratio like the above, of a coherent signal hypotheses $\mathcal{H}_S$ and an incoherent instrumental feature hypothesis $\mathcal{H}_I$ (the null-hypotheses) for a network of $D$ detectors. $\mathcal{H_I}$ states that each detector $i$ has either pure stationary Gaussian noise $\mathcal{H}_N$ or Gaussian noise and an incoherent noise transient (glitch) $\mathcal{H}_G$. Taking $Z^S$, $Z^G_i$ and $Z^N_i$ as the Bayesian evidences (defined in Appendix~\ref{sec:bayesianEvidEval}) for $\mathcal{H}_S$, $\mathcal{H}_N$, and $\mathcal{H}_G$, the \bcr is given by
\begin{equation}
\label{eq:bcr}
\bcr = \frac{P^S Z^S}{\prod\limits^D_{i=1} \ [P^G Z^G_i + (1-P^G)Z^N_i]}\ ,
\end{equation}
where $P^S$ and $P^G$ are the prior-odds of obtaining a signal or a glitch from a stretch of data. The prior-odds can be defined more explicitly as 
\begin{itemize}
    \item $P^S=P(\mathcal{H}_S)/P(\mathcal{H}_I)$, the prior-odds for obtaining a coherent signal versus an incoherent instrumental feature.
    \item $P^G=P(\mathcal{H}_G| \mathcal{H}_I)$, the prior-odds for obtaining a glitch assuming there is an incoherent instrumental feature.
\end{itemize}

When using the \bcr as a detection statistic, the physical interpretation of the prior-odds is lost. Hence, the prior-odds are empirically tuned to maximize the separation between the \bcr distribution of the background (expected to favor the $\mathcal{H}_I$ hypothesis) and the \bcr distribution of artificially manufactured simulated signals (expected to favor the $\mathcal{H}_S$ hypothesis). Increasing the separation between the two distributions can improve ability of the \bcr to discriminate candidate events as coherent signals or incoherent instrumental features. The tuning process is described in detail in Appendix~\ref{sec:tuning-prior-odds}.

\subsection{Estimation of astrophysical signal probability}
Candidate \bcr-values are either statistically insignificant compared to the background \bcr distribution, implying the candidate is more probable to be an incoherent instrumental feature (the $\mathcal{H}_I$ null-hypothesis), or statistically significant to the background distribution, indicating the possible presence of an astrophysical signal (the $\mathcal{H}_S$ hypothesis). A false alarm probability with trial factors, \fap, for the candidate \bcr-value can quantify the significance. The \fap is the probability that a candidate originating from a non-astrophysical source can be falsely identified as a signal.

To calculate the \fap, each candidate \bcr is considered a single statistical trial that can occur at a fixed false alarm probability $f$, where $f$ is the probability of observing a background $\bcr'$ greater than or equal to the candidate \bcr,
\begin{equation}
    f = \frac{\text{Count of } \bcr' \leq \bcr}{\text{Count of } \bcr'} \ .
\end{equation}
The false alarm probability with trials \fap that the \bcr measurement occurs at least once for $N$ trials $(N > 0)$, where $N$ is the number of candidate triggers is
\begin{equation}
    \fap = 1 - (1-f)^N \ .
\end{equation}

\avi{
Talk about the differences in pastro calculation ~\cite{Farr:2015:PhRvD, Gaebel:2019:MNRAS,Galaudage:2020:PhRvD}
}

\greg{This needs more explanation and discussion. I know that in the Gaebel paper~\cite{Gaebel:2019:MNRAS}, they make the connection (they don't in FGMC~\cite{Farr:2015:PhRvD}). But, effectively the FAP is the p-value, the probability that the event arose from the background noise alone. Taking 1-FAP as identifying a real signal can be hugely problematic (Wikipedia has its own page on it \url{https://en.wikipedia.org/wiki/Misuse_of_p-values}). In this case, it is a reasonable thing to do, because you are saying "anything which rings up our detection statistic is astrophysical by virtue that it is coherent". But, that needs to be spelled out. Also, it may be useful to say that the p-astro calculated from, say, PyCBC/GSTLAL are not 1-FAP, but instead apply the FGMC population approach.}

\subsection{Data for Analysis}
The LIGO Scientific collaboration operates several search pipelines that scan for gravitational-waves from compact binary mergers such as \gstlal~\cite{sachdev2019gstlal}, \texttt{MBTAOnline}~\cite{MBTA}, \texttt{SPIIR}~\cite{spiir} and \pycbc~\cite{pycbc_og4}. The output of \pycbc's search is a list of times and their corresponding \pycbc ranking statistic \pycbcstat values. The \pycbcstat ranking-statistic is akin to the matched-filter signal-to-noise ratio \snr. However, unlike \snr, \pycbcstat includes candidate signal's intrinsic and extrinsic properties and other information that feeds into determining if the signal can have astrophysical origins~\cite{pycbc_og6}. Whenever a local maximum of $\pycbcstat > \snr_\text{T}$, where $\snr_\text{T}$ is some predetermined threshold value, the \pycbc search pipeline produces a single-detector \textit{trigger} associated with the detector and time $t_c$ where the apparent signal in the data has its merger~\cite{pycbc_og6}.

When \pycbc observes a trigger between detectors with coincident parameters and a time of arrival difference less than the gravitational-wave travel time between detectors, the trigger is labeled a \textit{candidate event trigger}, a trigger that may be from astrophysical origins~\cite{pycbc_og1}. To test the pipeline's sensitivity \pycbc also conducts searches for \textit{simulated triggers}, artificial triggers manufactured by injecting signals into the detector data. Finally, to quantify the statistical significance of candidate triggers, \pycbc artificially constructs \textit{background triggers} to compare against the candidate events. These background triggers are coherent signal-free events, constructed by applying relative offsets, or time-slides, between the data of different detectors~\cite{pycbc_og6}. The background trigger's \pycbcstat distribution is used to calculate the candidate trigger's significance,  using null-hypothesis significance testing, under the assumption that all candidate event triggers are due to noise.

\subsection{Acquisition of triggers}
Advanced LIGO's second observing run O2 lasted $38$ weeks~\cite{GWOSC}. The software package, \pycbc~\cite{pycbc_code}, was used by LVK to process the O2 data in 22 time-frames (approximately 2 weeks per frame) and found several gravitational-wave events and numerous gravitational-wave candidates~\cite{pycbc_og0, pycbc_og1, pycbc_og2, pycbc_og3, pycbc_og4, pycbc_og5, pycbc_og6}. Some candidate events were vetoed to be glitches, while others were rejected due to their low significance. The data are divided into these time-frames because the detector's sensitivity does not stay constant throughout the eight-month-long observing period.

For our study, we filter the \pycbc background, simulated and candidate triggers to include only high-mass triggers with masses in the ranges of the parameters presented in Table~\ref{tab:parameters}. A plot of the \pycbc triggers from one time-frame, during April 23 - May 8, 2017, is presented in Fig.~\ref{fig:templateBank}. This figure also depicts the gravitational-wave templates used during the search through this time-frame of data.

}
% time for chunk 14
% 1176955218, Apr 23, 2017, 4:00 UTC
% 1178294418, May 08, 2017, 16:00 UTC
\caption[High-mass BCR search space.]{The template bank (pink) used by \pycbc to search a section of O2 data from $\text{April 23 - May 8, 2017}$. Our search is constrained to the high-mass parameter space enclosed by the dashed line. The candidate, background and simulated triggers detected in this region of the parameter space during this period are plotted in orange, gray and blue respectively.}\label{fig:templateBank}
\end{figure*}

\subsection{Calculating the BCR for triggers}
To evaluate $Z^S$, $Z^G_i$ and $Z^N_i$ and calculate the \bcr Eq.~\ref{eq:bcr} for triggers, we carry out Bayesian inference with \bilby~\cite{bilby, bilby_pipe}, employing \dynesty~\cite{dynesty} as our nested sampler. Nested sampling, an algorithm introduced by~\citet{skilling2004, skilling2006}, provides an estimate of the true Bayesian evidence and is often utilized for parameter estimation within the LIGO collaboration~\cite{bilby, bilby_paper, pbilby_paper}.

\begin{table}
    \centering
    \caption{
    Prior settings for the parameters used during our parameter estimation. The definitions of the parameters are documented in \citet{bilby_gwtc}~Table~E1.\label{tab:priors}}. The trigger time $t_c$ is obtained from the data products of \pycbc's O2 search.
    \begin{tabular}{c c c c}
    \hline
    Parameter & Shape & Limits \\
    \hline
          $\mathcal{M}/\msun$           & Uniform & 7--180  \\
          $q$                           & Uniform & 0.1--1  \\
          $M/\msun$                     & Constraint & 50--500  \\
          $d_\mathrm{L}/\mathrm{Mpc}$   & Comoving & 100--5000  \\
          $a_1$, $a_2$                  & Uniform & 0--1  \\
          $\theta_{JN}$                 & Sinusoidal & 0--$\pi$  \\
          $\psi$                        & Uniform & 0--$\pi$  \\
          $\phi$                        & Uniform & 0--$2\pi$  \\
          ra                            & Uniform & 0--$2\pi$  \\
          dec                           & Cosine & 0--$2\pi$  \\
          $t_c/\mathrm{s}$              & Uniform & $t_c\pm0.1$  \\
    \hline
    \end{tabular}
\end{table}

We take 31 neighboring, off-source, non-overlapping, 4-second segments of time-series data before the analysis data segment $d_i$ to generate the PSD. We use off-source to avoid the inclusion of a signal in the PSD calculation. A Tukey window with 0.2-second roll-offs is applied to each data segment to suppress spectral leakage after which the segments are fast-Fourier transformed and median-averaged to create a PSD~\cite{ligo_psd}. Like other PSD estimation methods, this method adds statistical uncertainties to the PSD~\cite{psd_student_t, chatziioannou2019noise}. To marginalize over the statistical uncertainty, we use the median-likelihood presented by~\citet{psd_student_t} as a post-processing step.

We take 31 neighboring off-source non-overlapping  4-second  segments of time-series data before the analysis data segment $d_i$ to generate the PSD.  A Tukey window with a 0.2-second roll-off is applied to each data segment to suppress spectral leakage. After this  the segments are fast-Fourier transformed and median-averaged to create a PSD~\cite{ligo_psd}. Like other PSD estimation methods, this method adds statistical uncertainties to the PSD~\cite{psd_student_t, chatziioannou2019noise, Biscoveanu:2020:PhRvD}. To marginalize over the statistical uncertainty, we use the median-likelihood presented by~\citet{psd_student_t} as a post-processing step. We find that this post-processing step improves the search efficiency by $49.26\%$ the details of this calculation are presented in the Appendix~\ref{sec:psd-marginalization}.

Finally, we neglect detector calibration uncertainty and acquire data from from the gravitational-wave Open Science Center~\cite{GWOSC}. The data we use are the publicly accessible O2 strain data from the Hanford and Livingston detectors, recorded while the detectors are in ``Science Mode''. We obtain the data using \gwpy~\cite{gwpy}.

\subsection{Assigning \pastro to candidate events}
After the calculating the \bcr for the entire set of high-mass background and simulated triggers, we calculate probability distributions $p_b(\bcr)$ and $p_s(\bcr)$ for each 2-week time-frame of O2 data. These distributions are used to `tune' prior-odd $P^S$ and $P^G$ values.

\begin{figure}[!ht]
{\centering \includegraphics[width=0.85\linewidth]{images/reweighted_bcr_cdf_smaller_legend.png} }
\caption[BCR distribution example]{Histograms represent the survival function (1-CDF) from our high-mass selection of background triggers (gray) and simulated signals (blue) triggers obtained from \pycbc's search of data from $\text{August 13 - 21, 2017}$. Vertical lines mark the $\text{ln}\ \bcr$ of \IAS's GW170817A and \GWTC's GW170814.}\label{fig:bcrCdf}
\end{figure}

\input{Data/results_table}
We analyze the $60,996$ background, $5,146$ simulated, and $25$ candidate triggers reported by \pycbc's search on the data from LIGO's second observing run, restricting our analysis to the triggers that fall within our mass-space as described in Section~\ref{sec:method}.We also analyze events and candidate events reported by \GWTC and the IAS group (note: some of these were identified as candidates by the \pycbc search). In Table~\ref{tab:results}, we summaries the \pastrobcr, along with the \pastro of other pipelines for comparison. Although the various pipeline \pastro are not mathematically equivalent, by comparing pipeline \pastro values for a given candidate, we can compare how significant each pipeline deems various candidates. The $P^S$ and $P^G$ values utilized for each time-frame are reported in Appendix~\ref{apdx:alphabeta}.

\subsection{GWTC-1 Events}
All the confirmed gravitational-wave events from binary black hole mergers reported in \GWTC and within our prior space, (specifically GW170104, GW170608, GW170729, GW170809 and GW170814), have $\pastrobcr>0.9$, indicating a high probability of the presence of an astrophysical signal.

In addition to the above confirmed gravitational-wave events from \GWTC, we have also analyzed several candidate events from \GWTC, most of which have low \pastrobcr. For example, consider the candidate event 170412, assigned a $\pastro$ of $0.06$ by \gstlal and has a $\pastrobcr$ of $0.01$. This candidate was reported to be excess power caused due noise appearing non-stationary between 60-200 Hz~\cite{GWTC1}. This candidate acts as an example of how $\pastrobcr$ may be utilized to eliminate candidates originating from terrestrial noise sources.

\subsection{IAS Events}
Our analysis of the high-mass IAS events and candidates in O2 has resulted in three events with disfavored $\pastrobcr<0.5$ (GWC170402, GW170403, GW170425), and four events and one candidate with $\pastrobcr\geq0.5$ (GW170121, 170302, GW170304, GW170727, GW170817A). While GW170727 and GW170817A's \pastrobcr are similar to the \pastro reported from IAS (the differences between the \pastro from \bcr and IAS is $|\Delta \pastro|<0.1$), the remaining candidates have opposing \pastro values (with $|\Delta \pastro| > 0.15$).
